---
title: "Installation"
date: "Last modified: `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5
)
```


This sections describes how to install the `{mapme.classification}` package. 
Because the package handles some spatial data, we assume that you have
a working installation of [GDAL](https://gdal.org/index.html) and [PROJ](https://proj.org/) 
on your system. Please refer to the respective installation guides to learn about how to install these programms
on you OS.

Windows users additionally will have to install the [rtools](https://cran.r-project.org/bin/windows/Rtools) 
software for a successful installation of the `{mapme.vegetation}` package from source. 

The installation of the `{mapme.classification}` package is then achieved trough the following
function call:


```{r installation, eval = F}
remotes::install_github("mapme-initiative/mapme.classification")
```

The installation process will install potential missing R dependencies.
After the successfull installation, you are ready to use the package's 
functionality. Simply load the package for your R session with the library
call:

```{r setup, val = F}
library(mapme.classification)
```

Note, that in order to use specific model algorithms via the `{caret}` package,
you might need to install certain system requirements manually. If that is the 
case you will be informed which packages to install when calling the modeling function.
Refer to the column `Library` to learn the name of the R package for a certain
algorithm under this [link](https://topepo.github.io/caret/available-models.html).
