---
title: "Workflow"
date: "Last modified: `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: mapme.classification.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.height = 5
)
```

## Workflow

In the following, the most important steps of the image classification workflow 
of `{mapme.classification}` are described; the figure guides the reader through these
descriptions. The goal of this tutorial is to classify Sentinel-2 images into 
different LULC categories in an agricultural landscape in Ethiopia, by using the 
`{mapme.classification}` package. Creating such LULC maps in different years would 
allow – for example – to back trace changes in the cropped area, which might be 
an intended outcome of a project being evaluated. Several algorithms, such as 
the Random Forest algorithm, can be calibrated based on user provided input data 
and reference data. The calibrated model is applied to classify the input data into user defined categories of land use / land cover (LULC). The assesses the accuracy of the LULC maps and estimates the area of each LULC category. Optimally, this LULC area estimation can be bias corrected by considering the class prevalence in the training data. The single processing steps are as follows:

-  pre-process Sentinel-2 images via the <code> {[mapme.vegetation](https://mapme-initiative.github.io/mapme.vegetation/}</code>
package
-  extract reference data for known locations
-  split reference data into training and validation set
-  training a model possibly based on spatio-temporal folds and a Forward-Feature-Selection
-  run the final model to yield spatially comprehensive predictions
-  optionally run post-processing routine to remove isolated pixels
-  return evaluation metrics of the final model based on test data
-  calculate LULC area and optionally correct for biases by considering the class prevalence in the training data

```{r, echo = FALSE}
library(DiagrammeR)

mermaid("
  graph TB

    classDef default fill:#c2dddd,stroke-width:2px,stroke:#034858;
 
    A(Extract reference data)
    B(Train/Test split)
    C(Calibrate a model)
    D(Spatial prediction)
    E(Post-processing)
    F(Accuracy assessment)
    G(Bias adjustment)
    A-->B
    B-->C
    C-->D
    D-->E
    D-->F
    D-->G

")
```

